---
title: "Project 2B"
author: "Vanita Thompson"
date: "3/4/2020"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
## Objective

The goal of this assignment is to give you practice in preparing different datasets for downstream
analysis work. I am using data on the cost of some popular WW2 airplanes. Since cost typically declines, 
I hope to show a decrease in the costs of airplanes overtime. 

## Load the Libraries

```{r setup, message=FALSE, warning=FALSE, comment = NA}
library(dplyr)
library(tidyr)
library(magrittr)
library(stringr)
library(zoo)
require(ggplot2)
require(ggthemes)
library(extrafont)
``` 

## Data Import

I copied this file from the web site linked below, and created a CSV file. I uploaded the file to github and I will read the untidy data into R. 
https://www.ibiblio.org/hyperwar/AAF/StatDigest/aafsd-3.html


```{r load, message=FALSE, warning=FALSE, comment = NA}
untidy= read.csv(file="https://raw.githubusercontent.com/Vthomps000/DATA607_VT/master/AirplaneCost.csv",header=TRUE,sep="\t",na.strings=c("","NA"))
head(untidy)
```

## Data Transformation

The data is in a poor format. There is missing/blank data. There is inconsistant notation, and empty cells. The currency symbol is used irregularly. I will tidy up this data and reorganize it with tidyr and dplyr.

```{r}
untidy$Type=na.locf(untidy$Type)
untidy$Type = str_trim(untidy$Type)
untidy$Model = str_trim(untidy$Model)
untidy$Type=as.factor(untidy$Type)
untidy$Model= as.factor(untidy$Model)
untidy = gather(untidy,Year,Cost,X1941:X1945)
untidy$Model[untidy$Model==""] <- NA
untidy$Model= as.factor(untidy$Model)
untidy$Year=str_replace(untidy$Year,"X","")
untidy$Year=as.factor(untidy$Year)
untidy$Cost=str_replace(untidy$Cost,",","")
untidy$Cost=as.numeric(untidy$Cost)
untidy = untidy[complete.cases(untidy$Model),]
write.csv(untidy,"Tidy_AirplaneCost.csv", row.names=FALSE)
head(untidy,6)
untidy
```

## Data Analysis 

I am comparing the data to show the decrease in cost of some fighter and bomber airplanes

```{r graphic, message=FALSE, warning=FALSE}
target=c("P-38","P-40","P-47","P-51")
fighter=filter(untidy,Model %in%target)   #filtering fighter planes
g=ggplot(data=fighter,aes(x=Year,y=Cost*0.001,group = Model, color = Model))
g=g+geom_line(stat="identity",size=1.6)
g=g+ggtitle("Cost of Fighters")
g=g+ylab("Cost ($K)")+xlab("Year")
g=g+theme_get()
g=g+theme(plot.title = element_text(hjust = 0.5),text=element_text(size=13, family="Times"))
g
```


Now I will compare the data with the Bomber planes.

```{r graphic2,message=FALSE, warning=FALSE}
target=c("B-17","B-24", "B-25", "B-26")
bomber=filter(untidy,Model %in%target)   #filtering bomber planes
g=ggplot(data=na.omit(bomber),aes(x=Year,y=Cost*0.001,group = Model, color = Model))
g=g+geom_line(stat="identity",size=1.6)
g=g+ggtitle("Cost of Bombers")
g=g+ylab("Cost ($K)")+xlab("Year")
g=g+theme_get()
g=g+theme(plot.title = element_text(hjust = 0.5),text=element_text(size=13, family="Times"))
g
```

## Conclusion

The data and graph confirm my original hypothesis. The cost of the airplanes dramatically decreased over time.