---
title: "Project 1"
author: "Vanita Thompson"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605
1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and
dividing by the total number of games played.
If you have questions about the meaning of the data or the results, please post them on the discussion forum. Data
science, like chess, is a game of back and forth…
The chess rating system (invented by a Minnesota statistician named Arpad Elo) has been used in many other contexts,
including assessing relative strength of employment candidates by human resource departments.
You may substitute another text file (or set of text files, or data scraped from web pages) of similar or greater
complexity, and create your own assignment and solution. You may work in a small team. All of your code should be in
an R markdown file (and published to rpubs.com); with your data accessible for the person running the script.


Required Data: Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

## Load the libraries 
```{r}
library(stringr)
library(DT)
library(data.table)
```

## Load the txt file

```{r}
tournament <- readLines("https://raw.githubusercontent.com/Vthomps000/DATA607_VT/master/tournamentinfo.txt")
head(tournament)
```

From observation of the data table, we see that the text file is formatted in a way where the information is segmented into 2 lines. The first line contains player ID, name and points with win/loss information. The second line contains state and rating. I will first format the row data into 2 different sets, 1 set for each row. From there, I will extract the necessary information to a vector. Finally, I will combined these vectors to form a formatted database.

## Subsetting the required data

```{r}
summary(tournament)
```

## Extracting relevant data by row: 

```{r}
name <- seq(5, 196, 3) 
state_point <- seq(6, 196, 3) 
```

## Extracting player ID and points from name

```{r}

player <- as.integer(str_extract(tournament[name], "\\d+")) 

name_player <- str_replace_all(str_extract(tournament[name],"([|]).+?\\1"),"[|]","") 

points <- str_extract(tournament[name], "\\d.\\d") 
```

## Extracting State and rating

```{r}
state <- str_extract(tournament[state_point], "[A-Z]{2}" ) 
rating <- as.integer(str_replace_all(str_extract(tournament[state_point], "R: \\s?\\d{3,4}"), "R:\\s", ""))
```



## combining the vectors to form the database

```{r}
new_tournament <- data.frame(player, name_player, state, points, rating) 
head(new_tournament)
```


## Calculating average opponent rating:

```{r}
opponent <- str_extract_all(str_extract_all(tournament[name], "\\d+\\|"), "\\d+")
avg_opp <- length(name)
for (i in 1:length(name)) 
{ 
  avg_opp[i] <- round(mean(rating[as.numeric(unlist(opponent[player[i]]))]), digits = 0)
}
avg_opp
```

## Adding the avg variable to the new database. I named this database newtournamentinfo

```{r}
newtournamentinfo <- data.frame(player, name, state, points, rating, avg_opp)
head(newtournamentinfo)
```

## Writing CSV

```{r}
write.csv(newtournamentinfo, file = "newtournamentinfo.csv")
```