---
title: "DATA 607 Final Project"
author: "Vanita Thompson"
date: "5/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Total COVID-19 Cases vs Population

## Introduction

In this project I will analyze the relationship between the Novel Coronavirus (COVID-19)and population around the world, to demonstrate dataprocessing and visualisation with R, tidyverse and ggplot2.

## Loading Libraries

```{r}
library(magrittr) # pipe operations
library(tidyverse) # ggplot2, tidyr, dplyr...
library(gridExtra) # multiple grid-based plots on a page
library(ggforce) # accelerating ggplot2
library(kableExtra) # complex tables
library(leaflet) # map
library(ggrepel)
```

## Data

The data source used for this analysis is from Our World in Data github repository found here:
<link> https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-data.csv 


## Load Data in R

At first, the dataset, which is a CSV files, is downloaded and saved as local files and then are loaded
into R from github.

```{r}
url <- 'https://raw.githubusercontent.com/Vthomps000/DATA607_VT/Final-Project/owid-covid-data.csv'
COVID19 <- read.csv(url)
c19 <- data.frame(COVID19)
head(c19)
```

```{r}
glimpse(c19)
```

The dataset has 171013 rows and 29 columns shown above.

## Data Cleaning

For this analysis I will be specifically looking at `location`,`total_cases`, and `population`.

```{r}
c19_2<- c19 %>%
  select(location, total_cases, population) %>%
  arrange(location)
```

The csv file entries were configured by date, causing duplicate entries for `location`. I collapsed all row entries by the number of `total_cases`.


```{r}
c19_3 <- count(c19_2, location, wt= total_cases)
names(c19_3)[names(c19_3) == "n"] <- "cases"
```

## Merging Datasets

Here, I merged the dataframes to regain the `population` column. Next I removed and replaced, `total_cases` with the column `cases` I created above.

```{r}
c19_4 <- merge(c19_2,c19_3, all.x = TRUE, all.y = TRUE)
```

```{r}
c19_5 <- c19_4 %>%
  select(-total_cases)
```

Here, I removed all the duplicate entries and row names.

```{r}
df_new <- unique(c19_5)
```

```{r}
rownames(df_new) <- c()
```

Now, I have a tidied dataset to work with.

```{r}
kable(df_new) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
 
## Data Analysis and Visualization

I wanted to take a look at the trends of all the available countries in the data set. However, that contained a large number of rows, so I decided to rank the top 20 countries by both `cases`, and `population` for comparison.


```{r}
ggplot(df_new, aes(x= log(population) , y= log(cases), label=location))+
  geom_point() +geom_text(aes(label=location),hjust=0, vjust=0, size=4)
```

```{r}
linearMod <- lm(population ~ cases, data=df_new)  # build linear regression model on full data
print(linearMod)
```

```{r}
cor.test(df_new$population, df_new$cases)
```

If the p Value is less than 0.05 and the cor test reult is 0.9404415, we reject the null hypothesis that the true correlation is zero (i.e. they are independent). Here, we reject the null hypothesis and conclude that cases is dependent on population.

```{r}
AIC(linearMod)
BIC(linearMod)
```
The Akaike’s information criterion - AIC (Akaike, 1974) and the Bayesian information criterion - BIC (Schwarz, 1978) are measures of the goodness of fit of an estimated statistical model and can also be used for model selection. Both criteria depend on the maximized value of the likelihood function L for the estimated model. Here, they are both quite high, indicating a poor fit.

```{r}
summary(linearMod)
```

Based on the linear model, the slope is positive, indicating a relationship between the `population` and number of `cases`. The $R^2$ is 0.8844, adjusted $R^2$ is 0.8839, the p value is far less than 0.5. However, there are many outliers in both variables.

```{r}
outlier_values <- boxplot.stats(df_new$cases)$out
outlier_values
```

```{r}
outlier_values <- boxplot.stats(df_new$population)$out
outlier_values
```


## Top 20 Countries Ranked

```{r}
## ranking by confirmed cases
df_20 <- df_new %>% 
select(location, population, cases) %>%
mutate(ranking = dense_rank(desc(cases)))
k <- 20
## top 20 countries: 21 incl. 'World'
top.countries <- df_20 %>% filter(ranking <= k + 1) %>%
arrange(ranking)
```

```{r}
kable(top.countries) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


```{r}
## ranking by population
df2_20 <- df_new %>% 
select(location, population, cases) %>%
mutate(ranking = dense_rank(desc(population)))
k <- 20
## top 20 countries: 21 incl. 'World'
top.countriespop <- df2_20 %>% filter(ranking <= k + 1) %>%
arrange(ranking)
```

```{r}
kable(top.countriespop) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```



```{r}
ggplot(top.countries, aes(x= log(population) , y= log(cases), label=location))+
  geom_point() +geom_text(aes(label=location),hjust=0, vjust=0, size=4)
```

```{r}
ggplot(top.countriespop, aes(x= log(population) , y= log(cases), label=location))+
  geom_point() +geom_text(aes(label=location),hjust=0, vjust=0, size=4)
```



```{r}
ggplot()+
  geom_point(data = top.countries, aes(x = log(population), y = log(cases), color = "blue"))+
  geom_text(data = top.countries, aes(x = log(population), y = log(cases), label= location, color = "blue"))+
  geom_point(data = top.countriespop, aes(x = log(population), y = log(cases), color = "red"), show.legend = TRUE)+
  geom_text(data = top.countriespop, aes(x = log(population), y = log(cases), label= location, color = "red"))+
  scale_color_manual(values = c("blue","red"), labels = c("top.countries","top.countriespop"), name = "")
```

## Conclusion

Based on the graphs and tables above there does appear to be some relationship between the population and the number of worlwide COVID-19 cases. However, this is just an observational study, as many variables were removed. Due to several other factors, including pre-existing health conditions and age, etc., we cannot confirm this hypothesis.
