---
title: "Project 2C"
author: "Vanita Thompson"
date: "3/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Assignment Description

Our task is to read in a .csv file and tidy the dataset. 

## Libraries

Load required libraries
```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(tibble)
library(ggplot2)
library(scales)
library(kableExtra)
```

## Data Import

I copied this file from the web site linked below, and created a CSV file. I uploaded the file to github and I will read the untidy data into R. 
https://towardsdatascience.com/whats-tidy-data-how-to-organize-messy-datasets-in-python-with-melt-and-pivotable-functions-5d52daa996c9

Read the .csv file from a github link
```{r message=FALSE, warning=FALSE}
url <- "https://raw.githubusercontent.com/Vthomps000/DATA607_VT/master/religion.csv"
untidy <- read_csv(url)
untidy <- as.tibble(untidy)
untidy
```

## Data Evaluation

Began by renaming an unecessary column
```{r}
clean <- untidy %>% rename(" " = X1)
clean
```

Reshaped the data

```{r}
cleaner <- clean %>% gather(`<10k`:`refused`, key = income, value = counts) %>% rename("Frequency" = counts) %>% select(religion, income, Frequency) %>%
arrange(religion)

cleaner
```



## Data Analysis & Visualization

I started by visualizing the whole data set. I found that most of the survey was conducted by Catholics so I decided to focus on them.

```{r}
target=c("<10k","10-20k","20-30k","30-40k", "40-50k", "50-75k", "75-100k", "100-150k", ">150k", "refused")
religion=filter(cleaner,income %in%target)   #filtering fighter planes
g=ggplot(data=religion,aes(x=income,y=Frequency ,group = religion, color = religion))
g=g+geom_line(stat="identity",size=1.6)
g=g+ggtitle("Income by religion")
g=g+ylab("# of People")+xlab("Income")
g=g+theme_get()
g=g+theme(plot.title = element_text(hjust = 0.5),text=element_text(size=13))
g
```

```{r}
religion2<- cleaner %>% select(religion, income, Frequency) %>%
  filter(religion == "Catholic")
religion2
```

```{r}
summary(religion2)
```




## Conclusion

I was abe to tidy the data set.


