---
title: "DATA 607 Homework 1"
author: "Vanita Thompson"
date: "2/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This week 1 assignment will subset data provided by FiveThirtyEight located here:

https://fivethirtyeight.com/features/what-do-men-think-it-means-to-be-a-man/

The dataset can be found here:

https://github.com/fivethirtyeight/data/tree/master/masculinity-survey

Author's at FiveThirtyEight conducted a survey with over 1600 men on whether or not the #MeTOO
movenment changed their view of masculinity.

##Load Library

```{r}
library(plyr)
```


## Load Data

Load masculinity data.

```{r}
x <- 'https://raw.githubusercontent.com/fivethirtyeight/data/master/masculinity-survey/raw-responses.csv'
y <- data.frame(read.csv(url(x), header = FALSE))
```

## Prepare Data

The data was subset to examine the salaries of heterosexual males that always pay when they take 
women out on dates.

```{r}
Men <- y[, c(61,84,86,88,87,95)]
colnames(Men) <- c("Pay","Orientation","Education","Salary","Location","Age")
```

## Display Results

Display subsetted column.

```{r}
subset(Men, Pay == "Always" & Orientation == "Straight", select = c("Salary"))
```

