---
title: "DATA607 HW 5"
author: "Vanita Thompson"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Tidying and Transforming Data

## Loading Libraries

```{r}
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(scales)
```


## CSV File

Data was entered in a csv file called thomps_HW5.csv

## Reading Data in R 

    CSV read into R from github.
    
```{r}
url <- "https://raw.githubusercontent.com/Vthomps000/DATA607_VT/master/thomps_HW5.csv"
dfuntidy <- read_csv(url)
dfuntidy
tbuntidy<-tbl_df(dfuntidy)
tbuntidy
```

## Data Transformation 

    We transform the data by first reshaping a wide table to a tidy table with variables
    as columns and observations as rows.

```{r}
tblTidy<-gather(dfuntidy,"Destinations","Flights",3:7)
tblTidy
```

    Next, we identify blanks and select all rows without blank values only, rename columns, 
    fill in missing info, create and match a new destination column with counts, and split 
    the status column.

```{r}
blanks <- seq(3,nrow(dfuntidy),3)
clean <- dfuntidy %>% slice(-blanks) %>% rename("airline" = X1, "status" = X2)
for (i in seq(from=2, to=nrow(clean), by=2)) {
  clean[i,1] <- clean[i-1,1]
}  
cleaner <- clean %>% gather(`Los Angeles`:`Seattle`, key = destination, value = counts)
tidy <- cleaner %>% spread(status, counts)
```

    The data is now tidy. Every column is a variable, where each row represents an observation,
    and every value has a cell.

## Analysis of Flight Arrivals and Delays

```{r}
tidy
```

    Using the tidy table with groups we can compare both airlines arrival times. The table 
    shows that AM West has more delayed flights than Alaska, but also has more on time. Since
    AM West has more flights than Alaska, we can compare is using ratios. Using pipes and mutate,
    We add a column of total flights and arrange them in descending order

```{r}
flights <- tidy %>% mutate(totalflights = delayed + `on time`) %>% arrange(desc(totalflights, 'on time', delayed))
flights
```

    To compare arrival ratios we can mutate our final table again to add a ratio column 
    for delayed flights (dratio) and on time flights (otratio)

```{r}
flights2 <- flights %>% mutate(dratio = delayed/totalflights, otratio = 1-dratio)
flights2
```

    From the ratio shown above we discover that Alaska has more delayed flights, with 14.2% of 
    its flights being delayed against 7.9% of AM West flights being delayed. We can also confirm
    Alaska has more on time flights overall by grouping and descending the on time ratios by destination.

```{r}
ot_flights <- flights2 %>% 
  arrange(destination, desc(otratio)) %>% 
  group_by(destination)  %>% 
  select(-delayed:-dratio)

ot_flights
```

##  Conclusion

    The ggplot graph supports our analysis of flight arrival times. Flights with Alaska
    face less delays than flights with AM West.

```{r}
ggplot(data=ot_flights, aes(x=reorder(destination, -otratio), y=otratio, fill=airline)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(labels = percent, breaks=seq(0,1,.1)) +
  scale_fill_brewer(palette="Blues") + theme_minimal() +
  labs(title="% of on time flights by destination", y = " % On Time Flights ", x = "Destination") 
```

