---
title: "DATA607 HW3"
author: "Vanita Thompson"
date: "2/12/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

##Load data

```{r}
majors <- read.csv(url('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv'), stringsAsFactors = F)
str(majors)
```

## Majors containing data or statistics:

```{r}
majors$Major[grepl("DATA", majors$Major)]
```

```{r}
majors$Major[grepl("STATISTICS", majors$Major)]
```

#2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

## Enter list values

```{r}
fruit_list <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'
```

##Load library

```{r}
library(stringr)
```

## unlist data

```{r}
foods <- str_extract_all(fruit_list, '[a-z]+\\s[a-z]+|[a-z]+')
unlist(foods)
```

#3 Describe, in words, what these expressions will match:
(.)\1\1

Any character repeated 3 times in a row 

```{r}
test_vec <- c("aaa", "aaba" , "abcdef" , "thbbbht")
nums <- unlist(str_extract_all(test_vec, "(.)\\1\\1"))
nums
```


“(.)(.)\\2\\1”

Two characters followed by the first two in reverse order

```{r} 
test_vec2 <- c("aabbaa", "aaba" , "abcdef")
nums2 <- unlist(str_extract_all(test_vec2, "(.)(.)\\2\\1"))
nums2
```

(..)\1

Two characters repeated twice.

```{r}
test_vec3 <- c("dodo", "ketchup" , "banana" , "cookie")
nums3 <- unlist(str_extract_all(test_vec3, "(..)\\1"))
nums3
```


“(.).\\1.\\1”

A character repeated three times with characters in between each repitition, e.g. abaca

```{r}
test_vec4 <- c("avocado", "cadaea" , "banana" , "cookie")
nums4 <- unlist(str_extract_all(test_vec4, "(.).\\1.\\1"))
nums4
```


"(.)(.)(.).*\\3\\2\\1"

The characters followed by any character repeate 0 or more times and then then the same three characters in reverse order.

```{r}
test_vec5 <- c("abc312131cba", "pizza", "bananaracecar" , "cookie")
nums5 <- unlist(str_extract_all(test_vec5, "(.)(.)(.).*\\3\\2\\1"))
nums5
```

#4 Construct regular expressions to match words that:

Start and end with the same character.

^(.).*\\1$

```{r}
test_vec6 <- c("abc312131cba", "pizza", "icecream" , "cookie")
nums6 <- str_extract(test_vec6, '^(.).*\\1$')
unlist(nums6)
```

 
Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)

(..).*\\1

```{r}
test_vec7 <- c("church", "bus", "river" , "mississippi")
nums7 <- str_extract_all(test_vec7, '(..).*\\1')
unlist(nums7)
```


Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)

([a-z]).*\\1.*\\1

```{r}
test_vec8 <- c("church", "eleven", "river" , "mississippi")
nums8 <- str_extract_all(test_vec8, '([a-z]).*\\1.*\\1')
unlist(nums8)
```

