---
title: "DATA605 HW3"
author: "Vanita Thompson"
date: "2/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

##Load data

```{r}
majors <- read.csv(url('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv'), stringsAsFactors = F)
str(majors)
```

## Majors containing data or statistics:

```{r}
majors$Major[grepl("DATA", majors$Major)]
```

```{r}
majors$Major[grepl("STATISTICS", majors$Major)]
```

#2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

## Enter list values

```{r}
fruit_list <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'
```

##Load library

```{r}
library(stringr)
```

## unlist data

```{r}
foods <- str_extract_all(fruit_list, '[a-z]+\\s[a-z]+|[a-z]+')
unlist(foods)
```

#3 Describe, in words, what these expressions will match:
(.)\1\1

The string ‘\1\1’ and the character before it, as long as it isn't a new line

“(.)(.)\\2\\1”

This could match anything like a 4 letter palindrome. one character, another character, the same as the 2nd character, the same as the first character, while surrounded by quotes.

(..)\1

This matches any two characters followed by the string ‘\1’

“(.).\\1.\\1”

This matches any character, followed by any other character, then the first character again, then any other character, then the first character again, while surrounded by quotes.

"(.)(.)(.).*\\3\\2\\1"

This would match any three characters and any or no characters in between the first 3 characters reversed, while surrounded by quotes.

#4 Construct regular expressions to match words that:

Start and end with the same character.

(.)[a-z]*\\1

Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)

([a-z]{2})[a-z]*\\1

Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)

[a-z]*([a-z])[a-z]*\\1[a-z]*\\1[a-z]*