---
title: "DATA 607 HW2"
author: "Vanita Thompson"
date: "2/8/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Running this markdown

1. Download the movies_data.sql file found in the Github repository and save it in your directory: https://github.com/Vthomps000/DATA607_VT/blob/master/movies_data.sql
2. Start MySQLWorkbench and load the script movies_data.sql
3. Set the movies schema as the default schema and run the script
4. Run code below replacing user and password with your own credentials
5. remove eval=FALSE from the chunk option

## Loading required Libraries

```{r}
library(RMySQL)
library(RMariaDB)
```

## Connecting to database server

Connecting to the MySQL server:

```{r include=FALSE}
# Replace the below with your credentials
user = 'root'
password = 'password'
```

## Loading the movie data

Load the movies data in the dataframe 'movies';

```{r eval=FALSE}
mydb = dbConnect(MySQL(), user = user, password = password, dbname = 'movies_data', host = 'localhost')
movies <- dbGetQuery(mydb, "select * from movies")
```
## Displaying data

There are 36 observations in this data set, 6 movies and 6 reviewers. Some values are NULL since not every participant has seen all of the movies.

```{r eval=FALSE}
summary(movies)
movies
```

## Omitting null values

Omit the NA values from the data:

```{r eval=FALSE}
movies <- na.omit(movies)
```

## Computing the Average

```{r warning=FALSE, eval=FALSE}
movies_avg_sql <- dbGetQuery(mydb, "select title, avg(rating) from movies group by title")
movies_avg_sql
```

Mean sorted by title:

```{r eval=FALSE}
aggregate(movies[, 6], list(movies$title), mean)
```