---
title: "Untitled"
author: "Vanita Thompson"
date: "3/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggthemes)
```


## CSV Data

Data was entered from https://www.usclimatedata.com/ in a csv file called ClimateData.csv using the wide format given

## Read CSV into R, tidy and transform data

CSV read into R

```{r}
dfWide<-read.csv(file="https://raw.githubusercontent.com/Vthomps000/DATA607_VT/master/ClimateData.csv")
dfWide
tbWide<-tbl_df(dfWide)
tbWide
```

## Data Transformation

We reshape the data by going from a wide table to a tidy table with variables as columns and observations as rows, and renaming the x column.

```{r}
clean <- dfWide %>% rename("Climate" = X)

clean
```

I transformed the data by organizing the 5 columns by month (as rows) to better visualize the climate data.

```{r}
cleaner <- clean %>% gather(`Jan`:`Dec`, key = Month , value = counts)
```

```{r}
tidy <- cleaner %>% spread(Climate, counts)
tidy
```

```{r}
tidy$Month <- factor(tidy$Month, month.abb, ordered=TRUE)
tidy[order(tidy$Month), ]
```

```{r}
summary(tidy)
```


## Data Analysis and Visualization

From the summary, there appears to be a relationship between average precipittion and hours of sunshine. 
I grouped those variables and used ggplot to confirm.

```{r}
tidy2 <- tidy %>% group_by(Month, `Av. precipitation in inch` ) %>% summarise(`Hours of sunshine`)

tidy2
```

```{r}
g=ggplot(tidy2, aes(x=`Av. precipitation in inch`,y=`Hours of sunshine`,group = Month, color = Month))
g=g+geom_point(stat="identity",size=8)
g=g+ggtitle("Precipitation Due to Sunshine")
g=g+ylab("Hours of Sunshine")+xlab("Av. precipitation in inch")
g=g+theme_get()
g=g+theme(plot.title = element_text(hjust = 0.5),text=element_text(size=13))
g  

g
```



