---
title: "DATA607 Tidyverse"
author: "Vanita Thompson"
date: "3/26/2020"
output:
  html_document: 
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  md_document:
      variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidyverse vignette with stringr


The `stringr` package  is the package of choice for working with character strings in Tidyverse.

We will be using the pipe notation `%>%`for the vignette.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

Load the data into R from Kaggle containing information about breweries in the United States from the following link: https://www.kaggle.com/brkurzawa/us-breweries/data .

```{r}
url <- 'https://raw.githubusercontent.com/Vthomps000/DATA607_VT/master/beer.csv'
beer <- read.csv(url)
head(beer)
```

# Detecting Matches

Detecing pattern matches in a string using str_detect, which returns boolean values.

```{r}
b0 <- beer$address %>% str_detect("PO")
head(b0)
```

```{r}
head(beer[b0,])
```

## Detecting matches using str_which

```{r}
b1 <- beer$address %>% str_which("PO")
head(b1)
```

We obtain the same result as above.

```{r}
head(beer[b1,])
```

## Mutating Strings

Modifying/mutating strings.

```{r}
inc <- beer$brewery_name %>% str_which("INC.|inc.|Inc.|INC|inc|Inc")
head(inc)
```


```{r}
t1 <- beer %>% select(brewery_name) %>% slice(inc)
head(t1)
```

```{r}
beer$brewery_name <- beer$brewery_name %>% str_replace("INC.|inc.|Inc.|INC|inc|Inc", "Inc.")
t2 <- beer %>% select(brewery_name) %>% slice(inc)
head(t2)
```

```{r}
t4 <- sapply(beer$address, function(x) str_to_upper(x))
head(t4)
```

## Subseting Strings


```{r}
t5 <- beer$address %>% str_extract_all("([:digit:]{5})$")
head(t5)
```

Joining and splitting strings.

```{r}
t6 <- beer$address %>% str_split(",")
head(t6)
```

I concentrate individual strings into one single string.

```{r}
t7 <- sapply(t6, function(x) str_c(x, collapse = ","))
head(t7)
```

## Conclusion

The `stringr` package of Tidyverse was used to demonstarte the manipulation of  strings.
 
## Citations
 
Information came from,  the `stringr` cheatsheet on [github](https://github.com/rstudio/cheatsheets/blob/master/strings.pdf).
